### 2016-08-24 doc/publications/plos16/plos16.tex

The digital revolution is seeing a dramatic increase in data collected about every aspect of 
life~\cite{Loebbecke2015}.  These data sets are not only growing vertically by capturing more and more events but also
horizontally by capturing more information about these events.  The challenge of big and `wide' data is especially
pronounced in the health space where whole genome sequence (WGS) technology enabled researchers to interrogate all 3
billion basepairs of the human genome.  While a large fraction of this data is identical between individuals, \marginpar{base-pairs}
identifying the relevant and disease specific differences is the focus of complex trait analysis [citation].

>>>>

The digital revolution is seeing a dramatic increase in data collected about every aspect of 
life~\cite{Loebbecke2015}.  These data sets are not only growing vertically by capturing more and more events but also
horizontally by capturing more information about these events.  The challenge of big and `wide' data is especially
pronounced in the health space where whole genome sequence (WGS) technology enabled researchers to interrogate all 3
billion base-pairs of the human genome.  
%While a large fraction of this data is identical between individuals,

Identifying the relevant and disease specific base-pair differences between individuals is the focus of genome wide association studies (GWAs)~\cite{CCC2007}, which typically ascertain the mutations prevalence between affected and unaffected individuals at every genomic position independently.  
% complex trait analysis~\cite{Yang2011}.
These calculations were typically on a reduced-representation of the genome (SNP-arrays), however with WGS becoming more prevalent deep GWAs  sample numbers Crop ~\cite{Yano2016}
